On the Synchronization Power of the Openflow Data Plane

### 2015-03-22 inband-STN.tex
completed section 4. need to update main pseudo code.

%\usepackage{framed}
\usepackage{framed,color}
\definecolor{shadecolor}{rgb}{0.9,0.9,0.9}

@@

\newcommand{\hide}[1]{}
\newcommand{\concat}[0]{\oplus}
\newcommand{\append}[0]{\frown}
\newcommand{\Nat}{\mathbb{N}}
\newcommand{\cas}{CAS}

\newcommand{\stefan}[1]{\textit{\textcolor{red}{[stefan]: #1}}} % Comments
\newcommand{\liron}[1]{\textit{\textcolor{green}{[liron]: #1}}} % Comments
\newcommand{\petr}[1]{\textit{\textcolor{blue}{[carlos]: #1}}} % Comments

@@

@@

We assume transactional access to  two kinds of objects: a
\emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
given (bounded or unbounded) set, and a read-write \emph{memory} abstraction
object, which can be used for storing the current used policy identifier.
%
\subsubsection{A resource claimer}
A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
\emph{unclaim} and \emph{abortIfClaimed} with the following sequential
semantics:
%The container synchronization object introduced in this paper
%is essentially a transactional memory abstraction which can be implemented
%in standard Openflow switches.
%Before presenting the technical details in the next section,%
%we give a brief overview.

@@

\item $\textit{abortIfClaimed}(k)$, $k\in\Nat$, 
abort current transaction if $k$ is currently claimed by any controller.
%returns $\textit{true}$ if and
%  only if $k$ is not currently claimed by any controller; if
%  $\textit{false}$ is returned, we say that the check operation
%  \emph{fails} (which causes an abort of the transaction invoking it).

@@

\item $\textit{abortIfUnequal}(addr,k)$, $addr,k\in\Nat$, 
aborts current transaction if the value at memory offset $addr$ is not equal to $k$.
%returns $\textit{true}$ if and
%  only if $k$ is not currently claimed by any controller; if
%  $\textit{false}$ is returned, we say that the check operation
%  \emph{fails} (which causes an abort of the transaction invoking it).

@@

%features , including
%\emph{bundling}, \emph{conditional application}, and \emph{groups}.
%[[PK let us leave for the moment
%We will describe later how to implement some of these features
%in older Openflow versions where they are not
%supported. \textbf{TODO}
%]]
Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of  
transactional operations  $\textit{claim}(x)$, $\textit{check}(x)$, $\textit{unclaim}(x)$,
and $\textit{CAS}(x)$, respectively. 
Algorithm~\ref{alg:commit} describes the procedure executed by
a transaction at \emph{commit time}.

@@

  \liron{got till here}

We use creation and deletion of flow entries as abstraction to claiming/unclaiming  resource id numbers and also to checking whether a given id is claimed. The match part of the added flow entry represents the added
number and the mask part of the match varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry. 
As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, the match mask is a concatenation of the controller id with all ones, while the value is the id concatenated twice. We can see that for any two numbers $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_1=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both claim and unclaim without affecting each other.

@@

@@

@@

\begin{algorithm}[t]
    \caption{Advanced policy update}
    \label{alg:update}
    \begin{algorithmic}[1]
    \Require update requirements $U$
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State \textbf{read} switch configuration
 			\State $my\_id\gets$ an unused \emph{id} 
 			\State $pid\gets$ current policy id 
 			\State $claim(pid)$
 			\State \textbf{read} switch configuration
 			\If {$pid\notin Cur$}
	 			\State $unclaim(pid)$
 				\State {\bf continue} (\textbf{restart} loop)
 			\EndIf
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTxn
 				\State $check(my\_id)$
	 			\State $CAS(pid,my\_id)$
	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTxn
	 		\State $unclaim(pid)$
     		%\State $res \gets $ commit transaction
     	\Until{$res\neq\texttt{abort}$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

@@

>>>>

%\usepackage{framed}
\usepackage{framed,color}
\definecolor{shadecolor}{rgb}{0.9,0.9,0.9}
\definecolor{mygreen}{rgb}{0,0.5,0}

@@

\newcommand{\hide}[1]{}
\newcommand{\concat}[0]{\oplus}
\newcommand{\append}[0]{\frown}
\newcommand{\Nat}{\mathbb{N}}
\newcommand{\cas}{CAS}
\newcommand{\claimcheck}{check}
\newcommand{\compare}{compare}

\newcommand{\stefan}[1]{\textit{\textcolor{red}{[stefan]: #1}}} % Comments
\newcommand{\liron}[1]{\textit{\textcolor{mygreen}{[liron]: #1}}} % Comments
\newcommand{\petr}[1]{\textit{\textcolor{blue}{[carlos]: #1}}} % Comments

@@

\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
\subsection{The flow entry modification command (FlowMod)}

\subsubsection{The CHECK\_OVERLAP flag}

\subsection{Bundling}

@@

We assume transactional access to  two kinds of objects: a
\emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
given (bounded or unbounded) set, and a read-write \emph{memory} abstraction
object, which can be used for storing the current used policy identifier.
%
\subsubsection{A resource claimer}
A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
\emph{unclaim} and \emph{\claimcheck} with the following sequential
semantics:
%The container synchronization object introduced in this paper
%is essentially a transactional memory abstraction which can be implemented
%in standard Openflow switches.
%Before presenting the technical details in the next section,%
%we give a brief overview.

@@

\item $\textit{\claimcheck}(k)$, $k\in\Nat$, 
abort current transaction if $k$ is currently claimed by any controller.
%returns $\textit{true}$ if and
%  only if $k$ is not currently claimed by any controller; if
%  $\textit{false}$ is returned, we say that the check operation
%  \emph{fails} (which causes an abort of the transaction invoking it).

@@

\item $\textit{\compare}(addr,k)$, $addr,k\in\Nat$, 
aborts current transaction if the value at memory offset $addr$ is not equal to $k$.
%returns $\textit{true}$ if and
%  only if $k$ is not currently claimed by any controller; if
%  $\textit{false}$ is returned, we say that the check operation
%  \emph{fails} (which causes an abort of the transaction invoking it).

@@

%features , including
%\emph{bundling}, \emph{conditional application}, and \emph{groups}.
%[[PK let us leave for the moment
%We will describe later how to implement some of these features
%in older Openflow versions where they are not
%supported. \textbf{TODO}
%]]
Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of  
transactional operations  $\textit{claim}(x)$, $\textit{\claimcheck}(x)$, $\textit{unclaim}(x)$, $\textit{write}(x)$,  $\textit{\compare}(x)$
and $\textit{\cas}(x)$, respectively. 
Algorithm~\ref{alg:commit} describes the procedure executed by
a transaction at \emph{commit time}.

@@

 As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, for the resource-claimer, 
  the match part of the added flow entry represents the claimed id (id concatenated twice).
   In addition the mask part of the match (controller id concatenated with all-ones) varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry. 
    We can see that for any two resource ids $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_1=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both claim and unclaim without affecting each other.

In order to check if a number is claimed we try to create the corresponding entry while setting the
flag \textsf{CHECK\_OVERLAP} and using different action, thereby inflicting a failure in case an entry with overlapping match value exists. If the check succeeds we delete the entry. Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$. Therefore if one is used in check after the other used in claims, a conflict is detected iff $x_1=x_2$, as expected.
Note that checking for claims must contain a deletion of the added flow entry otherwise a successive check operation might mistakenly fail. Moreover the two commands (adding and deleting of the flow entries) must be sent as part of the same bundle to support concurrent checks by different controllers.

@@

\begin{algorithm}[t]
    \caption{$\textit{\claimcheck}(x)$}
    \label{alg:check}
    \begin{algorithmic}[1]
    \Require \emph{self} as the calling controller id.
%    \Ensure
    		\State $value \gets x\concat x$
    		\State $mask \gets self\concat all-1$
    		\State $match \gets (value,mask)$
    		%\State $cookie \gets x\concat self$
    		\State $action \gets x\concat self$
    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
			\Return $cmd1$,$cmd2$
    \end{algorithmic}
\end{algorithm}

@@

We use similar techniques for implementing memory write and abortIfUnequal operations.
 As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare}, 
  the match part of the added flow entry is used as the memory address and the action as the written value. Values written to the same address will replace one another, following the command definition to replace old flow entry with a new one in casae they share the same match parts (as described in Section \ref{sec:background}).

In order to check if a value is written in a givven address, we try to add it the same way as in the write operation  but we also set the \textsf{CHECK\_OVERLAP} flag. If the value is not there, our action differ from the action of the existsing entry thereby inflicting a failure.
In case the value is there, our flow entry replaces the existing flow entry with identical one.
In addition we present in Algorithm \ref{alg:cas} a simple implementation of Comapare and set (CAS) usin the two last operations.

\begin{algorithm}[t]
    \caption{$\textit{write}(addr,k)$}
    \label{alg:write}
    \begin{algorithmic}[1]
    %\Require  id $i$.
%    \Ensure
    		\State $value \gets addr$
    		\State $mask \gets  all-1$
    		\State $match \gets (value,mask)$
    		%\State $cookie \gets x\concat self$
    		\State $action \gets k$
    		\State $flag \gets 0$
    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{$\textit{\compare}(addr,k)$}
    \label{alg:compare}
    \begin{algorithmic}[1]
    %\Require  id $i$.
%    \Ensure
    		\State $value \gets addr$
    		\State $mask \gets  all-1$
    		\State $match \gets (value,mask)$
    		%\State $cookie \gets x\concat self$
    		\State $action \gets k$
    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{$\textit{\cas}(addr, old,new)$}
    \label{alg:cas}
    \begin{algorithmic}[1]
    %\Require  id $i$.
%    \Ensure

    		\State $cmd1\gets \textsc{abortIfUnequal}((addr, old) $
    		\State $cmd2\gets \textsc{Write}((addr, new) $
			\Return $cmd1,cmd2$
    \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[t]
    \caption{Advanced policy update}
    \label{alg:update}
    \begin{algorithmic}[1]
    \Require update requirements $U$
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State \textbf{read} switch configuration
 			\State $my\_id\gets$ an unused \emph{id} 
 			\State $pid\gets$ current policy id 
 			\State $claim(pid)$
 			\State \textbf{read} switch configuration
 			\If {$pid\notin Cur$}
	 			\State $unclaim(pid)$
 				\State {\bf continue} (\textbf{restart} loop)
 			\EndIf
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTxn
 				\State $abortIfClaimed(my\_id)$
	 			\State $CAS(pid,my\_id)$
	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTxn
	 		\State $unclaim(pid)$
     		%\State $res \gets $ commit transaction
     	\Until{$res\neq\texttt{abort}$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

@@

\section{Implementing CAS with GroupMod commands}\label{sec:with-groups}
Similarly, we use a deletion of group $x$ and a creation of group $y$ (with GroupMod commands) as an
abstraction to performing a $CAS(x,y)$ operation to a special $16$bit register (we abstract only one register). The $16$ bit limit is resulted by th $16$bit group id size. 
%The group id of the added group represents the added number. 
Note that trying to delete a group that doesn't exist (or trying to create a group with id that already exists) fails, thereby preventing adding the new value $y$ if $x$ hasn't existed before. Note that the two action should be performed inside the same bundle in order to ensure the atomic nature ot the abstraction.

\begin{algorithm}[H]
    \caption{$\textit{group-CAS}(old,new)$}
    \label{alg:gcas}
    \begin{algorithmic}[1]
    %\Require  id $i$.
%    \Ensure
